---
title             : "NFL Big Data Bowl"
subtitle          : "A Final Project for Data III"
author: 
  - name          : "Sean Xiaohao Duan"
    affiliation   : "1"
    email         : "sxdff5@mail.missouri.edu"
  - name          : "Ronald D. Flores"
    affiliation   : "1"
    email         : "ronald.flores@mail.missouri.edu"
  - name          : "Hyejin Shim"
    affiliation   : "1"
    email         : "hs3z8@mail.missouri.edu"
  
affiliation:
  - id            : "1"
    institution   : "University of Missouri"


authornote: |
  
  This paper was prepared as part of a submission for a Kaggle Project, as a final project for a Statistics seminar at the University of Missouri. All the authors participated in the seminar.

abstract: |
  PUT IN AN INTRO/ABSTRACT HERE
  
date: "12/17/2020"
output: html_document
---

```{r pkgload, include = FALSE}
#PUT YOUR LIBRARY PACKAGES HERE WHEN YOU ARE READY
library(lme4)
library(ggplot2)
library(papaja)
library(sjPlot)
library(papaja)
library(brms)
library(dplyr)
library(lubridate)
library(stringr)
```

# Introduction

We decided to choose the NFL Big Data Bowl as our Kaggle project because all three of our group members had a reasonable understanding and enjoyment of professional American football. Additionally, we were all very happy with the abundance of real-world data that required minimal finessing and cleaning to get it to an analysis ready position. Lastly, because of the variety and depth of the data available we believed that we would be able to find and answer many interesting questions.

```{r child = "david_primary.Rmd"}
```


```{r child = "sean_primary.Rmd"}
```


```{r child = "joe_primary.Rmd"}
```


```{r child = "discussion.Rmd"}
```

# Computational Details
All results were obtained using the R system for statistical
computing. Code for reproducing our results can be found at https://github.com/seanxduan/Data-III-Project-3.